Using Component Model to implement a DSL: The Multi-Stencil Language Case Study

### 2016-11-08 ARTICLES/OTHER/C2S_EXA/slides.tex
final

%-------------------------------------------------------------------
\title[The Multi-Stencil Language]{Using Component Model to implement a DSL: The Multi-Stencil Language Case Study}
\author[]{Hélène Coullon, Julien Bigot, \underline{Christian Perez}} % [Hélène Coullon (INRIA), Julien Bigot (CEA), Christian Perez (INRIA)]
\institute[]{Avalon project team\\Maison de la simulation}
\date{C2S@Exa -- Scientific Days \\Paris, November 8th, 2016}
%-------------------------------------------------------------------

@@

@@

%% \begin{center}
%% Separation of concerns between domain/parallelization and optimization\\
%% Domain specific languages (DSL)\\
%% \medskip
%% \resizebox{2cm}{!}{\includegraphics{images/scientist.jpg}}
%% \end{center}
%% \end{frame}
%-------------------------------------------------------------
\begin{frame}
  \frametitle{Domain Specific Languages} % HPC DSL
  \begin{block}{Advantages}
    \begin{itemize}
    \item Easy language for end users (can be application specific)
    \item Separation of concerns (domain/implementation)
    \item Implicit parallelization and optimizations
    \end{itemize}
  \end{block}
  \begin{alertblock}{Limitations}
    \begin{itemize}
    \item Difficulties deported to the DSL designer and implementer
      \begin{itemize}
      \item Low level high performance programming
      \item Maintainability and portability
      \end{itemize}
    \item Difficult to combine DSLs (interoperability)
      \begin{itemize}
      \item exascale applications = many specific domains and interactions
      \end{itemize}
    \end{itemize}
  \end{alertblock}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
  \frametitle{Component Models} % components
  \begin{block}{Overview}
    \begin{itemize}
    \item Component = black box with well defined interactions (provided/required)
    \item Application = Assembly of components
    \end{itemize}
  \end{block}
  \begin{block}{Benefits}
    \begin{itemize}
    \item Common model to define software architecture
    \item Maintainability through separation of concerns
    \item Code-reuse and productivity
    \end{itemize}
  \end{block}
  \begin{alertblock}{Open issues}
    \begin{itemize}
    \item Component for HPC, eg task graph?\\ {\color{red} Cf next talk of Jérôme Richard!}
    \end{itemize}
  \end{alertblock}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{This talk}
%implicitly parallel (one sep. of concerns) / specific / performance
\begin{block}{From a DSL: Multi Stencil Language}
\begin{itemize}
\item Descriptive language,
\item for Multi-Stencil simulations,
\item without numerical code.
\end{itemize}
\end{block}

@@

%-------------------------------------------------------------
\section{The Multi-Stencil Language}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: the mesh}
\begin{center}
\resizebox{6cm}{!}{\includegraphics{images/mesh.pdf}}
\end{center}
\begin{columns}
\begin{column}{6.5cm}
%    \fontsize{11}{11}\selectfont
  \begin{lstlisting}[frame=single]
mesh : cart
mesh entities : cell, edgex
computation domains :
  d1 in cell
  d2 in edgex
stencil shapes : 
  ncc from cell to cell
  nce from cell to edgex
  nec from edgex to cell
\end{lstlisting}
\end{column}
\begin{column}{5.5cm}
\begin{itemize}
\item A cartesian mesh is used
\item Two kinds of mesh entities are declared onto it
\item Two computations domains, one for each entity type
\item Three stencil shapes (neighborhood from mesh entity to mesh entity)
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: quantity}
\begin{columns}
\begin{column}{5cm}
\begin{lstlisting}[frame=single]
quantity :
  A,cell
  B,cell
  C,edgex
  D,cell
  E,cell
  F,cell
  G,cell
  H,edgex
  I,cell
  J,cell
scalar : mu, tau
\end{lstlisting}
\end{column}
\begin{column}{6cm}
\begin{itemize}
\item A is a quantity applied onto cell
\item C is a quantity applied onto edgex
\item mu and tau are scalar values
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: time and computations}
\begin{columns}
\begin{column}{6cm}
\begin{lstlisting}[frame=single]
time : 500
computations :
  B[d1] = k0({tau},{A})
  C[d2] = k1({},{B[nce]})
  D[d1] = k2({},{C})
  E[d1] = k3({},{C})
  F[d1] = k4({},{D,C[nec]})
  G[d1] = k0({mu,tau},{E})
  H[d2] = k6({},{F})
  I[d1] = k7({},{G,H})
  J[d1] = k8({mu},{I[ncc]})
\end{lstlisting}
\end{column}
\begin{column}{5cm}
\begin{itemize}
\item The time loop is composed of 500 iterations.
\item J is written
\item onto the computation domain d1,
\item by the computation k8,
\item which read the scalar mu
\item and the quantity I,
\item which is accessed with the neighborhood ncc.
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------

@@

\begin{center}
\textcolor{red}{Static} or \textcolor{red}{dynamic} scheduling ?
\end{center}
\end{frame}
%----------------
%----------------
\begin{frame}
\frametitle{MSL to Component-based runtime}
\begin{block}{Ready-to-fill parallel pattern}
\begin{itemize}
\item Data parallelism
\begin{itemize}
\item External distributed data structure
\item Automatic detection of synchronizations
\end{itemize}
\item Task parallelism (mid-grain)
\begin{itemize}
\item Static scheduling at the assembly level 
\item Dynamic scheduling within a driver component
  \begin{itemize}
  \item OpenMP task (For loop also available)
  \end{itemize}
\end{itemize}
\end{itemize}
\end{block}
The fine grain task parallelism is left to other languages: 
\begin{itemize}
\item OpenMP in the kernels
\item Kernels generated by stencil compilers
  \begin{itemize}
  \item Pochoir, PATUS, Liszt etc.
  \end{itemize}
\end{itemize}
\end{frame}
%----------------
%-------------------------------------------------------------
\begin{frame}
\frametitle{Component Assembly Runtime Template}
A data parallel/task parallel component assembly.\vspace{-0.2em}

@@

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Cp) edge [solid] node {} (C)
      (C) edge [solid] node {} (Cu)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (DDSp2) edge [solid] node {} (DDS)
      (C)  edge [solid] node {} (Cu2)
          edge [solid] node {} (Cu3)
          edge [solid] node {} (Cu4)
      (Cu2)  edge [solid] node {} (k0p)
      (k0p)  edge [solid] node {} (k0)
      (Cu3)  edge [solid] node {} (k0sp)
      (k0sp)  edge [solid] node {} (k0s)
      (Cu4)  edge [solid] node {} (kotherp)
      (kotherp)  edge [solid] node {} (kother);
    \end{tikzpicture}
    };
    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=4cm of Data] (mpidata) {};
    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
    \node[solid,above=3.2cm of proc.west,anchor=west] (core) {\textit{\tiny{Applied on each resource}}};
    \path[-]
      (Data.east) edge node {} (mpidata);
  \end{tikzpicture}
\end{center}
{\fontsize{8}{8}\selectfont DDS: Distributed Data Structure (using SkelGIS)\par}
\end{frame}
%% %-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{What is produced ?}
%% \begin{block}{Implementation}
%% \begin{itemize}
%% \item DDS + Quantity + $k_{*}^{sync}$ = written components using SkelGIS (distributed memory, MPI)
%% \item Scheduler = Serie-Parallel graph of $\Gamma_{dep}$ (OpenMP 3.0)
%% \item Time + Scheduler = generated components
%% \item $k_i$ = empty components to write
%% \end{itemize}
%% \end{block}

%% \begin{block}{Separation of concerns}
%% \begin{itemize}
%% \item numerician (MSL)
%% \item computer engineer (K)
%% \item HPC engineer (DDS, quantity, $k_{*}^{sync}$ etc.)
%% \item DSL designer (component assembly + generated components)
%% \end{itemize}
%% \end{block}
%% \end{frame}
%% %-------------------------------------------------------------

%% \begin{frame}
%% \frametitle{Series-Parallel Tree}
%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
%% \normalsize
%% \begin{columns}
%% \begin{column}{.48\textwidth}
%% \begin{center}
%% \resizebox{\textwidth}{!}{%
%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};

%%    \node[] (c0)    at (-3,1) {$c_0$};
%%    \node[] (star1) at (-2,1) {$sync_1$};
%%    \node[] (c1)    at (-1,1) {$c_1$};
%%    \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
%%    \node[] (c7)    at ( 1,1) {$c_7$};
%%    \node[] (star8) at ( 2,1) {$sync_8$};
%%    \node[] (c8)    at ( 3,1) {$c_8$};

%%    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
%%    \node[] (s2)    at (1,2) {$\mathcal{S}$};

%%    \node[] (c2)    at (-3,3) {$c_2$};
%%    \node[] (star4) at (-2,3) {$sync_4$};
%%    \node[] (c4)    at (-1,3) {$c_4$};
%%    \node[] (c6)    at ( 0,3) {$c_6$};
%%    \node[] (c3)    at ( 1,3) {$c_3$};
%%    \node[] (c5)    at ( 2,3) {$c_5$};

%%   \path[->]
%%     (s0) edge node {} (c0)
%%          edge node {} (star1)
%%          edge node {} (c1)
%%          edge node {} (p0)
%%          edge node {} (c7)
%%          edge node {} (star8)
%%          edge node {} (c8)
%%     (p0) edge node {} (s1)
%%          edge node {} (s2)
%%     (s1) edge node {} (c2)
%%          edge node {} (star4)
%%          edge node {} (c4)
%%          edge node {} (c4)
%%          edge node {} (c6)
%%     (s2) edge node {} (c3)
%%          edge node {} (c5);
%%   \end{tikzpicture}
%%   }
%% \end{center}
%% \end{column}
%% \begin{column}{.52\textwidth}

%% \end{column}
%% \end{columns}
%% \end{frame}
%% %----------------

%% %----------------
%% \begin{frame}
%% \frametitle{Series-Parallel Tree}
%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
%% \normalsize
%% \begin{columns}
%% \begin{column}{.48\textwidth}
%% \begin{center}
%% \resizebox{\textwidth}{!}{%
%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};

%%    \node[] (c0)    at (-3,1) {$c_0$};
%%    \node[] (star1) at (-2,1) {$sync_1$};
%%    \node[] (c1)    at (-1,1) {$c_1$};
%%    \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
%%    \node[] (c7)    at ( 1,1) {$c_7$};
%%    \node[] (star8) at ( 2,1) {$sync_8$};
%%    \node[] (c8)    at ( 3,1) {$c_8$};

%%    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
%%    \node[] (s2)    at (1,2) {$\mathcal{S}$};

%%    \node[] (c2)    at (-3,3) {$c_2$};
%%    \node[] (star4) at (-2,3) {$sync_4$};
%%    \node[] (c4)    at (-1,3) {$c_4$};
%%    \node[] (c6)    at ( 0,3) {$c_6$};
%%    \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
%%    %\node[] (c5)    at ( 2,3) {$c_5$};

%%   \path[->]
%%     (s0) edge node {} (c0)
%%          edge node {} (star1)
%%          edge node {} (c1)
%%          edge node {} (p0)
%%          edge node {} (c7)
%%          edge node {} (star8)
%%          edge node {} (c8)
%%     (p0) edge node {} (s1)
%%          edge node {} (s2)
%%     (s1) edge node {} (c2)
%%          edge node {} (star4)
%%          edge node {} (c4)
%%          edge node {} (c4)
%%          edge node {} (c6)
%%     (s2) edge node {} (c3);
%%   \end{tikzpicture}
%%   }
%% \end{center}
%% \end{column}
%% \begin{column}{.52\textwidth}
%% \end{column}
%% \end{columns}
%% Loop fusion optimization possible
%% \end{frame}
%% %----------------

%% %----------------
%% \begin{frame}
%% \frametitle{Series-Parallel Tree}
%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
%% \normalsize
%% \begin{columns}
%% \begin{column}{.48\textwidth}
%% \begin{center}
%% \resizebox{\textwidth}{!}{%
%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};

>>>>

\AtBeginSection[]
{
\begin{frame}{Table of Contents}
\tableofcontents[currentsection]
\end{frame}
}
%-------------------------------------------------------------------
\title[The Multi-Stencil Language]{Using Component Model to implement a DSL: The Multi-Stencil Language Case Study}
\author[]{Hélène Coullon, Julien Bigot, \underline{Christian Perez}} % [Hélène Coullon (INRIA), Julien Bigot (CEA), Christian Perez (INRIA)]
\institute[]{Avalon project team\\Maison de la simulation}
\date{C2S@Exa -- Scientific Days \\Paris, November 8th, 2016}
%-------------------------------------------------------------------

@@

\begin{frame}{Table of Contents}
  \tableofcontents
\end{frame}

@@

%% \begin{center}
%% Separation of concerns between domain/parallelization and optimization\\
%% Domain specific languages (DSL)\\
%% \medskip
%% \resizebox{2cm}{!}{\includegraphics{images/scientist.jpg}}
%% \end{center}
%% \end{frame}
%-------------------------------------------------------------
\begin{frame}
  \frametitle{Domain Specific Languages} % HPC DSL
  \begin{block}{Advantages}
    \begin{itemize}
    \item Easy language for end users (can be application specific)
    \item Separation of concerns (domain/implementation)
    \item Implicit parallelization and optimizations
    \end{itemize}
  \end{block}
  \begin{alertblock}{Limitations}
    \begin{itemize}
    \item Difficulties deported to the DSL designer and implementer
      \begin{itemize}
      \item Low level high performance programming
      \item Maintainability and portability
      \end{itemize}
    \item Difficult to combine DSLs (interoperability)
      \begin{itemize}
      \item exascale applications = many specific domains and interactions
      \end{itemize}
    \end{itemize}
  \end{alertblock}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
  \frametitle{Component Models} % components
  \begin{block}{Overview}
    \begin{itemize}
    \item Technology advocating composition rather than development
      \begin{itemize}
      \item Old idea (late 60') but major development in the 90'
      \item After Object and before Model Driven Engineering (MDE)
      \end{itemize}
    \item Component = black box with well defined interactions
      \begin{itemize}
      \item Provides / Requires
      \end{itemize}
    \item Application = Assembly of components
    \end{itemize}
  \end{block}
  \begin{block}{Benefits}
    \begin{itemize}
      \item Application structure well defined
    \item Code-reuse and productivity
    \item Maintainability through separation of concerns
    \end{itemize}
  \end{block}
\vspace{-4cm}\hspace{8.15cm}\resizebox{2.5cm}{!}{\includegraphics{images/assembly.png}}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{L2C: A HPC Component Model}
\begin{block}{A minimalist component model for HPC}
  \begin{itemize}
  \item Component creation/deletion, configuration, and connection
  \item No L2C code between components \@ runtime
    \begin{itemize}
      \item No language interoperability provided
    \end{itemize}
  \item  Support of native interactions
    \begin{itemize}
    \item Use/Provide: C++, CORBA, [ FORTRAN (2008)]
    \item MPI communicator
    \end{itemize}
  \item \url{hlcm.gforge.inria.fr}
  \end{itemize}
\end{block}
\begin{alertblock}{Ongoing extensions}
  \begin{itemize}
  \item Concurrent reconfiguration
    \begin{itemize}
    \item DirectL2C
    \end{itemize}
  \item Task support
    \begin{itemize}
    \item {\color{red} Cf next talk on Comet!}
    \end{itemize}
  \end{itemize}
\end{alertblock}
\vspace{-4.2cm}\hspace{6.8cm}\resizebox{4cm}{!}{\includegraphics{images/jacobi-2X2.png}}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{This talk}
%implicitly parallel (one sep. of concerns) / specific / performance
\begin{block}{From a DSL: Multi Stencil Language}
\begin{itemize}
\item Descriptive language,
\item for Multi-Stencil simulations,
\item without numerical code.
\end{itemize}
\end{block}

@@

%-------------------------------------------------------------
\section{The Multi-Stencil Language}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: the mesh}
\begin{center}
\resizebox{6cm}{!}{\includegraphics{images/mesh.pdf}}
\end{center}
\begin{columns}
\begin{column}{6.5cm}
%    \fontsize{11}{11}\selectfont
  \begin{lstlisting}[frame=single]
mesh : cart
mesh entities : cell, edgex
computation domains :
  d1 in cell
  d2 in edgex
stencil shapes : 
  ncc from cell to cell
  nce from cell to edgex
  nec from edgex to cell
\end{lstlisting}
\end{column}
\begin{column}{5.5cm}
\begin{itemize}
\item A Cartesian mesh is used
\item Two kinds of mesh entities are declared onto it
\item Two computations domains, one for each entity type
\item Three stencil shapes (neighborhood from mesh entity to mesh entity)
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: quantity}
\begin{columns}
\begin{column}{5cm}
\begin{lstlisting}[frame=single]
quantity :
  A,cell
  B,cell
  C,edgex
  D,cell
  E,cell
  F,cell
  G,cell
  H,edgex
  I,cell
  J,cell
scalar : mu, tau
\end{lstlisting}
\end{column}
\begin{column}{6cm}
\begin{itemize}
\item A is a quantity applied onto cell
\item C is a quantity applied onto edgex
\item mu and tau are scalar values
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: time and computations}
\begin{columns}
\begin{column}{6cm}
\begin{lstlisting}[frame=single]
time : 500
computations :
  B[d1] = k0({tau},{A})
  C[d2] = k1({},{B[nce]})
  D[d1] = k2({},{C})
  E[d1] = k3({},{C})
  F[d1] = k4({},{D,C[nec]})
  G[d1] = k0({mu,tau},{E})
  H[d2] = k6({},{F})
  I[d1] = k7({},{G,H})
  J[d1] = k8({mu},{I[ncc]})
\end{lstlisting}
\end{column}
\begin{column}{5cm}
\begin{itemize}
\item The time loop is composed of 500 iterations.
\item J is written
\item onto the computation domain d1,
\item by the computation k8,
\item which read the scalar mu
\item and the quantity I,
\item which is accessed with the neighborhood ncc.
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------

@@

\begin{center}
\textcolor{red}{Static} or \textcolor{red}{dynamic} scheduling ?
\end{center}
\end{frame}
%----------------
%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
\normalsize
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

   \node[] (c0)    at (-3,1) {$c_0$};
   \node[] (star1) at (-2,1) {$sync_1$};
   \node[] (c1)    at (-1,1) {$c_1$};
   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
   \node[] (c7)    at ( 1,1) {$c_7$};
   \node[] (star8) at ( 2,1) {$sync_8$};
   \node[] (c8)    at ( 3,1) {$c_8$};

   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
   \node[] (s2)    at (1,2) {$\mathcal{S}$};

   \node[] (c2)    at (-3,3) {$c_2$};
   \node[] (star4) at (-2,3) {$sync_4$};
   \node[] (c4)    at (-1,3) {$c_4$};
   \node[] (c6)    at ( 0,3) {$c_6$};
   \node[] (c3)    at ( 1,3) {$c_3$};
   \node[] (c5)    at ( 2,3) {$c_5$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3)
         edge node {} (c5);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}

\end{column}
\end{columns}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
\normalsize
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

   \node[] (c0)    at (-3,1) {$c_0$};
   \node[] (star1) at (-2,1) {$sync_1$};
   \node[] (c1)    at (-1,1) {$c_1$};
   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
   \node[] (c7)    at ( 1,1) {$c_7$};
   \node[] (star8) at ( 2,1) {$sync_8$};
   \node[] (c8)    at ( 3,1) {$c_8$};

   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
   \node[] (s2)    at (1,2) {$\mathcal{S}$};

   \node[] (c2)    at (-3,3) {$c_2$};
   \node[] (star4) at (-2,3) {$sync_4$};
   \node[] (c4)    at (-1,3) {$c_4$};
   \node[] (c6)    at ( 0,3) {$c_6$};
   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
   %\node[] (c5)    at ( 2,3) {$c_5$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}
\end{column}
\end{columns}
Loop fusion optimization possible
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
\normalsize
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

   \node[] (c0)    at (-3,1) {$c_0$};
   \node[] (star1) at (-2,1) {$sync_1$};
   \node[] (c1)    at (-1,1) {$c_1$};
   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
   \node[] (c7)    at ( 1,1) {$c_7$};
   \node[] (star8) at ( 2,1) {$sync_8$};
   \node[] (c8)    at ( 3,1) {$c_8$};

   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
   \node[] (s2)    at (1,2) {$\mathcal{S}$};

   \node[] (c2)    at (-3,3) {$c_2$};
   \node[] (star4) at (-2,3) {$sync_4$};
   \node[] (c4)    at (-1,3) {$c_4$};
   \node[] (c6)    at ( 0,3) {$c_6$};
   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
   %\node[] (c5)    at ( 2,3) {$c_5$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}
\begin{block}{Specific components}
\begin{itemize}
\item $SEQ$ to directly replace $\mathcal{S}$ nodes
\item $PAR$ to directly replace $\mathcal{P}$ nodes
\item $SYNC$ for synchronizations
\item $K$ for computation kernels
\end{itemize}
\end{block}
\end{column}
\end{columns}
Loop fusion optimization possible
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{MSL to Component-based runtime}
\begin{block}{Ready-to-fill parallel pattern}
\begin{itemize}
\item Data parallelism
\begin{itemize}
\item External distributed data structure
\item Automatic detection of synchronizations
\end{itemize}
\item Task parallelism (mid-grain)
\begin{itemize}
\item Static scheduling at the assembly level 
\item Dynamic scheduling within a driver component
  \begin{itemize}
  \item OpenMP task
  \end{itemize}
\end{itemize}
\end{itemize}
\end{block}
The fine grain task parallelism is left to other languages: 
\begin{itemize}
\item OpenMP in the kernels
\item Kernels generated by stencil compilers
  \begin{itemize}
  \item Pochoir, PATUS, Liszt etc.
  \end{itemize}
\end{itemize}
\end{frame}
%----------------
%-------------------------------------------------------------
\begin{frame}
\frametitle{Component Assembly Runtime Template}
A data parallel/task parallel component assembly.\vspace{-0.2em}

@@

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Cp) edge [solid] node {} (C)
      (C) edge [solid] node {} (Cu)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (DDSp2) edge [solid] node {} (DDS)
      (C)  edge [solid] node {} (Cu2)
          edge [solid] node {} (Cu3)
          edge [solid] node {} (Cu4)
      (Cu2)  edge [solid] node {} (k0p)
      (k0p)  edge [solid] node {} (k0)
      (Cu3)  edge [solid] node {} (k0sp)
      (k0sp)  edge [solid] node {} (k0s)
      (Cu4)  edge [solid] node {} (kotherp)
      (kotherp)  edge [solid] node {} (kother);
    \end{tikzpicture}
    };
    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=4cm of Data] (mpidata) {};
    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
    \node[solid,above=3.2cm of proc.west,anchor=west] (core) {\textit{\tiny{Applied on each resource}}};
    \path[-]
      (Data.east) edge node {} (mpidata);
  \end{tikzpicture}
\end{center}
{\fontsize{8}{8}\selectfont DDS: Distributed Data Structure (using SkelGIS)\par}
\end{frame}
%----------------
\begin{frame}
\frametitle{Component-based runtime: Static Scheduling}

\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
     \node[component,solid] (D) at (0,0) {$Driver$};
     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
     %\node[solid] (Ds) at (-1.5,0) {start};
     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
     %\node[solid,left=0.8cm of T] (tt) {$T$};

     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
     \node[component,dashed,right=2cm of Cp] (C) {$Computations$};
     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
     \node[solid,right=0.2cm of Cu] (star) {$*$};
     %\node[solid,right=1.5cm of C] (gamma) {$\Gamma$};

     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
     %\node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};

     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
     %\node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
